%---------------- COMMENT FOR IMPORTING ----------------------
%\RequirePackage{lineno}				
%\documentclass[12pt]{report}		
%\pagestyle{headings}
%\input{ST-input2}							

%\setcounter{chapter}{0}
%\begin{document}								
%\setpagewiselinenumbers
%\linenumbers
%\tableofcontents
%-------------------------------------------------------------

\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

I would like to thank my supervisors Dr Nicos Pavlidis and Professor Idris Eckley for their endless support, guidance and patience throughout this project. Thanks are also due to Professor Lyudmila Mihaylova for her help with the conference paper presented in Chapter 4. The data analysed in Chapter 3 was kindly provided by Shell.

My PhD has been supported by the EPSRC funded Statistics and Operational Research (STOR-i) Centre for Doctoral Training. The STOR-i CDT has provided an excellent research environment and many additional training opportunities which have allowed me to develop my research skills.  Particular thanks go to the STOR-i director Jon Tawn for his constant support, especially during the late PhD stages. 

I would also like to give special thanks  to the `11-`15 cohort of STOR-i: Ben, Dave, Emma, Hugo, James, Judd, Rob and Tom. My PhD experience was greatly enhanced by their knowledge, encouragement and friendship.  

Finally, I would like to thank my family, Geraint, Joan, Sian and Tom. Dach chi'n werth y byd. 

%---------------- COMMENT FOR IMPORTING ----------------------
%\pagebreak											%Comment for importing
%\bibliographystyle{plainnat}		%Comment for importing
%\bibliography{References}				%Comment for importing
%\end{document}									%Comment for importing
%-------------------------------------------------------------
